animbook: Visualizing changes in performance measures and demographic affiliations using animation

### 2023-11-27 animbook-journal.Rmd
change the paper to the new package structure

In designing the \pkg{animbook} package, a three-step structured approach was developed to create an animation. The initial step is to reformat the data to the animated data structure as seen in Figure \@ref(fig:data-diagram). The second stage is creating a `ggplot` object, which can be subsequently passed into the animation function. The final step involves transforming the `ggplot` to a `gganimate` object by integrating the animation settings. This three-step structure was implemented to ensure that users, regardless of their level of experience, can produce the animations with simplicity while retaining customization for more experienced users.

@@

rank <- anim_prep(dbl_change, id, values, time, color = gp)

rank_data <- rank$data |> 
  filter(id %in% c(1, 14, 21, 100, 106, 148)) |> 
  rename(gp = color,
         rank = qtile) |> 
  select(-frame) |> 
  arrange(gp, id)

absolute <- anim_prep(dbl_change, id, values, time, color = gp, scaling = "absolute")

absolute_data <- absolute$data |> 
  filter(id %in% c(1, 14, 21, 100, 106, 148)) |>  
  rename(gp = color,
         absolute = qtile) |> 
  select(-frame) |> 
  arrange(gp, id)

rank_group <- anim_prep(dbl_change, id, values, time, color = gp, group_scaling = gp)

rank_group_data <- rank_group$data |> 
  filter(id %in% c(1, 14, 21, 100, 106, 148)) |> 
  rename(gp = group,
         rank_group = qtile) |> 
  select(-c(frame, color)) |> 
  arrange(gp, id)

absolute_group <- anim_prep(dbl_change, id, values, time, color = gp, group_scaling = gp, scaling = "absolute")

absolute_group_data <- absolute_group$data |> 
  filter(id %in% c(1, 14, 21, 100, 106, 148)) |> 
  rename(gp = group,
         absolute_group = qtile) |> 
  select(-c(frame, color)) |> 
  arrange(gp, id)
```

@@

  The frame variable is assigned by sorting the data based on the ID and time using the `arrange()` function, followed by applying the `group_by()` function on the ID, allowing the `row_number()` function to be performed within each group. The functions mentioned in this paragraph are from the \CRANpkg{dplyr} package (@dplyr).

```{r}
argument <- c("data", "id", "values", "time", "ngroup", "breaks", "group_scaling",
               "scaling", "order", "label", "color", "time_dependent", "runif_min",
               "runif_max")

description <- c("A data frame containing the data to be prepared for visualization.",
                 "The column name that represents the unique identifier variable.",
                 "The column name that contains the numeric values to be visualized.",
                 "The column name represents the time variable.",
                 "The number of groups or categories to create for scaling values.",
                 "A vector of breaks for creating bins.",
                 "The column name that represents the grouping variable.",
                 "The scaling method to be used; \"rank\" or \"absolute.\"",
                 "A vector of order for sorting the category values.",
                 "A vector of labels to be used for the y-axis in the visualization.",
                 "The column name to be used in ggplot2::aes() for the plot function.",
                 "Logical. Should the visualization be time-dependent? Default is TRUE.",
                 "The minimum value for random addition to frame numbers.",
                 "The maximum value for random addition to frame numbers.")

@@

```{r tbl-animprep}
argument_table |> 
  filter(argument %in% c("ngroup", "breaks", "group_scaling", "scaling")) |> 
  knitr::kable(caption = "The argument in the anim\\_prep for customized the data scaling.")
```

@@

```{r tbl-visual}
argument_table |> 
  filter(argument %in% c("label", "color", "time_dependent", 
                         "runif_min", "runif_max")) |> 
  knitr::kable(caption = "The additional argument for customizing the aesthetics of the visualization.")
```

@@

Once the data is prepared. The next step is to create the `ggplot` object as a basis for the animation. There are three plots available in this package. Two of the plots could be used for the animation, and another plot is used as a static visualization. All of the plots have an internal function that converts the standard data format into the required structure for each plotting function.

@@

  The `wallaby_data()` function is responsible for performing data manipulation and formatting tasks on the original object. It includes creating additional data components for labeling and shading. This function also responds by interpolating the non-linear path. It performs this task by mapping the `sine()` function provided in this package to each observation using the `map()` function from the \CRANpkg{purrr} package (@purrr). The frame is recalculated using the same method mentioned in the data preprocessing section. The `sankey_shade()` function is called to generate the proportional shaded data.

```{r shade-algorithm, fig.show="hold", out.width="50%", fig.cap="The plot shows how the algorithm for the sankey\\_shade() function works. The left figure represents the initial step of the algorithm, which calculates all the corner points. The right figure demonstrates the subsequent step, where points in-between the left and right are interpolated using the sine() function."}
knitr::include_graphics("figures/sankey-shade-1.png")
knitr::include_graphics("figures/sankey-shade-2.png")
```

  The algorithm behind the `sankey_shade()` function for generating the proportional shaded data is illustrated in Figure \@ref(fig:shade-algorithm). It started by calculating the corner points for all of the shaded areas. Then, use the `sine()` function to interpolate the point between the left and right.

@@

```{r, eval=knitr::is_html_output()}
argument2 <- c("object", "group_palette", "shade_palette", "rendering",
               "subset", "relation", "total_point", "height", "width", "size",
               "alpha")

description2 <- c("The animbook object returned from the prep function.",
                  "The vector of the palette used by the function to supply the color to each group.",
                  "The vector of the palette used by the function to supply the color to the shaded area.",
                  "The choice of method used to create and display the plot, either gganimate or plotly.",
                  "A character string specifying the variable used for subsetting the data. The \"top\" and \"bottom\" strings can also be used in this argument.",
                  "The choice of relationship for the values to display on the plot, either \"one_many.\" or \"many_one.\"",
                  "The number of points the users want for the wallaby plot. Default is NULL, which is the number of points equal to the original.",
                  "The proportion of the area occupied by the observations in the shaded areas.",
                  "The distance between the first and last observation in the animation.",
                  "The point size.",
                  "The opacity of the proportional shaded areas.")

@@

```{r, eval=knitr::is_latex_output()}
argument2 <- c("object", "group_palette", "shade_palette", "rendering",
               "subset", "relation", "total_point", "height", "width", "size",
               "alpha")

description2 <- c("The animbook object returned from the prep function.",
                  "The vector of the palette used by the function to supply the color to each group.",
                  "The vector of the palette used by the function to supply the color to the shaded area.",
                  "The choice of method used to create and display the plot, either gganimate or plotly.",
                  "A character string specifying the variable used for subsetting the data. The \"top\" and \"bottom\" strings can also be used in this argument.",
                  "The choice of relationship for the values to display on the plot, either \"one_many.\" or \"many_one.\"",
                  "The number of points the users want for the wallaby plot. Default is NULL, which is the number of points equal to the original.",
                  "The proportion of the area occupied by the observations in the shaded areas.",
                  "The distance between the first and last observation in the animation.",
                  "The point size.",
                  "The opacity of the proportional shaded areas.")

@@

In this section, the package usage will be demonstrated using the `cat_change` dataset included in this package. This dataset is simulated data that has changed from category A to E between two time points. Since this data is already in the categorized data structure, the `anim_prep_cat()` function is utilized.

  The output produced from this package is referred to as the `animbook` object. This contained a list of data and settings that hold prepared data for visualization, along with a list of settings used in the plot function.

```{r, echo=TRUE}
animbook <- anim_prep_cat(cat_change, 
                          id = id, 
                          values = qnt, 
                          time = time, 
                          color = gp, 
                          time_dependent = FALSE)

head(animbook$data, 10)
str(animbook)
```

@@

```{r, echo=TRUE, fig.width=8, fig.align='center', out.width="100%", layout = "l-body"}
p <- wallaby_plot(object = animbook,
             group_palette = RColorBrewer::brewer.pal(9, "Set1"),
             shade_palette = c("#737373", "#969696", "#BDBDBD",
                               "#D9D9D9","#D9D9D9","#D9D9D9"),
             rendering = "ggplot",
             subset = "top",
             relation = "one_many",
             total_point = NULL)

@@

gganimate::animate(p2)
```
```{r catchange-figure, fig.cap = "Animate visualization using example data. All of the X observations stay within the same group, while Y observations change from group A to group E.", fig.width=8, fig.align='center', out.width="100%", layout = "l-page", eval=knitr::is_latex_output()}
knitr::include_graphics("figures/animation-example.png")
```

@@

accounting <- anim_prep(data, 
                      id = ID, 
                      values = sales, 
                      time = year, 
                      label = label, 
                      ngroup = 4, 
                      color = country, 
                      time_dependent = FALSE)

@@

```{r, echo=TRUE}
voter <- anim_prep_cat(data = aeles,
                       id = id,
                       values = party,
                       time = year,
                       color = gender,
                       order = NULL,
                       time_dependent = FALSE)

p_voter <- wallaby_plot(object = voter,
                  group_palette = c("pink", "blue", "red"),
                  shade_palette = c("#737373", "#969696", "#BDBDBD",
                                    "#D9D9D9","#D9D9D9","#D9D9D9"),
                  rendering = "ggplot",
                  subset = "top",
                  relation = "one_many",
                  height = 1,
                  size = 2.5,
                  width = 100,
                  total_point = 1000)

>>>>

In designing the \pkg{animbook} package, a three-step structured approach was developed to create an animation. The initial step is to reformat the data into a categorized data structure, as seen in Figure \@ref(fig:data-diagram). The second stage is creating a `ggplot` object, which can be subsequently passed into the animation function. During the second stage, an internal function will turn the categorized data into an animated data structure. The final step involves transforming the `ggplot` to a `gganimate` object by integrating the animation settings. This three-step structure was implemented to ensure that users, regardless of their level of experience, can produce the animations with simplicity while retaining customization for more experienced users.

@@

rank <- anim_prep(dbl_change, id, values, time, group = gp)

class(rank) <- c("tbl_df", "tbl", "data.frame")

rank_data <- rank |> 
  filter(id %in% c(1, 14, 21, 100, 106, 148)) |> 
  rename(gp = group,
         rank = qtile) |>  
  arrange(gp, id)

absolute <- anim_prep(dbl_change, id, values, time, group = gp, scaling = "absolute")

class(absolute) <- c("tbl_df", "tbl", "data.frame")

absolute_data <- absolute |> 
  filter(id %in% c(1, 14, 21, 100, 106, 148)) |>  
  rename(gp = group,
         absolute = qtile) |> 
  arrange(gp, id)

rank_group <- anim_prep(dbl_change, id, values, time, group = gp, group_scaling = gp)

class(rank_group) <- c("tbl_df", "tbl", "data.frame")

rank_group_data <- rank_group |> 
  filter(id %in% c(1, 14, 21, 100, 106, 148)) |> 
  rename(gp = group,
         rank_group = qtile) |> 
  arrange(gp, id)

absolute_group <- anim_prep(dbl_change, id, values, time, group = gp, group_scaling = gp, scaling = "absolute")

class(absolute_group) <- c("tbl_df", "tbl", "data.frame")

absolute_group_data <- absolute_group |> 
  filter(id %in% c(1, 14, 21, 100, 106, 148)) |> 
  rename(gp = group,
         absolute_group = qtile) |> 
  arrange(gp, id)
```

@@

  The frame variable is assigned by sorting the data based on the ID and time using the `arrange()` function, followed by applying the `group_by()` function on the ID, allowing the `row_number()` function to be performed within each group. The functions mentioned in this paragraph are from the \CRANpkg{dplyr} package (@dplyr). These are done by the internal function for each of the plots provided in the packages.

```{r}
argument <- c("data", "id", "values", "time", "group", "ncat", "breaks", "label",
              "group_scaling", "scaling", "order", "time_dependent")

description <- c("A data frame containing the data to be prepared for visualization.",
                 "The column name that represents the unique identifier variable.",
                 "The column name that contains the numeric values to be visualized.",
                 "The column name represents the time variable.",
                 "The column name that distinguishes between the values group.",
                 "The number of groups or categories to create for scaling values.",
                 "A vector of breaks for creating bins.",
                 "A vector of labels to be used for the y-axis in the visualization.",
                 "The column name that represents the grouping variable.",
                 "The scaling method to be used; \"rank\" or \"absolute.\"",
                 "A vector of order for sorting the category values.",
                 "Logical. Should the visualization be time-dependent? Default is TRUE.")

@@

```{r tbl-animprep}
argument_table |> 
  filter(argument %in% c("ncat", "breaks", "group_scaling", "scaling")) |> 
  knitr::kable(caption = "The argument in the anim\\_prep for customized the data scaling.")
```

@@

```{r tbl-visual}
argument_table |> 
  filter(argument %in% c("label", "group", "time_dependent")) |> 
  knitr::kable(caption = "The additional argument for customizing the aesthetics of the visualization.")
```

@@

Once the data is prepared. The next step is to create the `ggplot` object as a basis for the animation. There are three plots available in this package. Two of the plots could be used for the animation, and another plot is used as a static visualization. All of the plots have an internal function that converts the categorized data format into the animated structure for each plotting function.

@@

  The `wallaby_data()` function is responsible for performing data manipulation and formatting tasks on the original object. It includes creating additional data components for labeling and shading. This function also responds by interpolating the non-linear path. It performs this task by mapping the `sine()` function provided in this package to each observation using the `map()` function from the \CRANpkg{purrr} package (@purrr). The frame is recalculated using the same method mentioned in the data preprocessing section. The `proportional_shade()` function is called to generate the proportional shaded data.

```{r shade-algorithm, fig.show="hold", out.width="50%", fig.cap="The plot shows how the algorithm for the proportional\\_shade() function works. The left figure represents the initial step of the algorithm, which calculates all the corner points. The right figure demonstrates the subsequent step, where points in-between the left and right are interpolated using the sine() function."}
knitr::include_graphics("figures/sankey-shade-1.png")
knitr::include_graphics("figures/sankey-shade-2.png")
```

  The algorithm behind the `proportional_shade()` function for generating the proportional shaded data is illustrated in Figure \@ref(fig:shade-algorithm). It started by calculating the corner points for all of the shaded areas. Then, use the `sine()` function to interpolate the point between the left and right.

@@

```{r, eval=knitr::is_html_output()}
argument2 <- c("data", "group_palette", "shade_palette", "rendering",
               "subset", "relation", "total_point", "height", "width", "size",
               "alpha")

description2 <- c("The categorized data returned from the prep function.",
                  "The vector of the palette used by the function to supply the color to each group.",
                  "The vector of the palette used by the function to supply the color to the shaded area.",
                  "The choice of method used to create and display the plot, either gganimate or plotly.",
                  "A character string specifying the variable used for subsetting the data. The \"top\" and \"bottom\" strings can also be used in this argument.",
                  "The choice of relationship for the values to display on the plot, either \"one_many.\" or \"many_one.\"",
                  "The number of points the users want for the wallaby plot. Default is NULL, which is the number of points equal to the original.",
                  "The proportion of the area occupied by the observations in the shaded areas.",
                  "The distance between the first and last observation in the animation.",
                  "The point size.",
                  "The opacity of the proportional shaded areas.")

@@

```{r, eval=knitr::is_latex_output()}
argument2 <- c("data", "group_palette", "shade_palette", "rendering",
               "subset", "relation", "total_point", "height", "width", "size",
               "alpha")

description2 <- c("The categorized data returned from the prep function.",
                  "The vector of the palette used by the function to supply the color to each group.",
                  "The vector of the palette used by the function to supply the color to the shaded area.",
                  "The choice of method used to create and display the plot, either gganimate or plotly.",
                  "A character string specifying the variable used for subsetting the data. The \"top\" and \"bottom\" strings can also be used in this argument.",
                  "The choice of relationship for the values to display on the plot, either \"one_many.\" or \"many_one.\"",
                  "The number of points the users want for the wallaby plot. Default is NULL, which is the number of points equal to the original.",
                  "The proportion of the area occupied by the observations in the shaded areas.",
                  "The distance between the first and last observation in the animation.",
                  "The point size.",
                  "The opacity of the proportional shaded areas.")

@@

In this section, the package usage will be demonstrated using the `cat_change` dataset included in this package. This dataset is simulated data that has changed from category A to E between two time points. Since this data is already in the categorized data structure, the `anim_prep_cat()` function is utilized. The output produced from this package is referred to as the `categorized` data.

```{r, echo=TRUE}
animbook <- anim_prep_cat(cat_change, 
                          id = id, 
                          values = qnt, 
                          time = time, 
                          group = gp)

head(animbook, 10)
str(animbook)
```

@@

```{r, echo=TRUE, fig.width=8, fig.align='center', out.width="100%", layout = "l-body"}
p <- wallaby_plot(data = animbook,
             group_palette = RColorBrewer::brewer.pal(9, "Set1"),
             shade_palette = c("#737373", "#969696", "#BDBDBD",
                               "#D9D9D9","#D9D9D9","#D9D9D9"),
             rendering = "gganimate",
             subset = "top",
             relation = "one_many",
             total_point = NULL)

@@

```{r catchange-figure, fig.cap = "Animate visualization using example data. All of the X observations stay within the same group, while Y observations change from group A to group E.", fig.width=8, fig.align='center', out.width="100%", layout = "l-page", eval=knitr::is_latex_output()}
knitr::include_graphics("figures/animation-example.png")
```

@@

accounting <- anim_prep(data, 
                      id = ID, 
                      values = sales, 
                      time = year, 
                      label = label, 
                      ncat = 4, 
                      group = country)

@@

```{r, echo=TRUE}
voter <- anim_prep_cat(data = aeles,
                       id = id,
                       values = party,
                       time = year,
                       group = gender,
                       order = NULL)

p_voter <- wallaby_plot(data = voter,
                  group_palette = c("pink", "blue", "red"),
                  shade_palette = c("#737373", "#969696", "#BDBDBD",
                                    "#D9D9D9","#D9D9D9","#D9D9D9"),
                  rendering = "gganimate",
                  subset = "top",
                  relation = "one_many",
                  height = 1,
                  size = 2.5,
                  width = 100,
                  total_point = 1000)

### 2023-11-27 animbook-journal.tex
change the paper to the new package structure

In designing the \pkg{animbook} package, a three-step structured approach was developed to create an animation. The initial step is to reformat the data to the animated data structure as seen in Figure \ref{fig:data-diagram}. The second stage is creating a \texttt{ggplot} object, which can be subsequently passed into the animation function. The final step involves transforming the \texttt{ggplot} to a \texttt{gganimate} object by integrating the animation settings. This three-step structure was implemented to ensure that users, regardless of their level of experience, can produce the animations with simplicity while retaining customization for more experienced users.

@@

The frame variable is assigned by sorting the data based on the ID and time using the \texttt{arrange()} function, followed by applying the \texttt{group\_by()} function on the ID, allowing the \texttt{row\_number()} function to be performed within each group. The functions mentioned in this paragraph are from the \CRANpkg{dplyr} package (Wickham et al. (2023)).

@@

\caption{\label{tab:tbl-animprep}The argument in the anim\_prep for customized the data scaling.}
\centering
\begin{tabular}[t]{l|l}
\hline
Argument & Description\\
\hline
ngroup & The number of groups or categories to create for scaling values.\\
\hline
breaks & A vector of breaks for creating bins.\\
\hline
group\_scaling & The column name that represents the grouping variable.\\
\hline
scaling & The scaling method to be used; "rank" or "absolute."\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:tbl-visual}The additional argument for customizing the aesthetics of the visualization.}
\centering
\begin{tabular}[t]{l|l}
\hline
Argument & Description\\
\hline
label & A vector of labels to be used for the y-axis in the visualization.\\
\hline
color & The column name to be used in ggplot2::aes() for the plot function.\\
\hline
time\_dependent & Logical. Should the visualization be time-dependent? Default is TRUE.\\
\hline
runif\_min & The minimum value for random addition to frame numbers.\\
\hline
runif\_max & The maximum value for random addition to frame numbers.\\
\hline
\end{tabular}
\end{table}

@@

Once the data is prepared. The next step is to create the \texttt{ggplot} object as a basis for the animation. There are three plots available in this package. Two of the plots could be used for the animation, and another plot is used as a static visualization. All of the plots have an internal function that converts the standard data format into the required structure for each plotting function.

@@

The \texttt{wallaby\_data()} function is responsible for performing data manipulation and formatting tasks on the original object. It includes creating additional data components for labeling and shading. This function also responds by interpolating the non-linear path. It performs this task by mapping the \texttt{sine()} function provided in this package to each observation using the \texttt{map()} function from the \CRANpkg{purrr} package (Wickham and Henry (2023)). The frame is recalculated using the same method mentioned in the data preprocessing section. The \texttt{sankey\_shade()} function is called to generate the proportional shaded data.

\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/sankey-shade-1} \includegraphics[width=0.5\linewidth]{figures/sankey-shade-2} \caption{The plot shows how the algorithm for the sankey\_shade() function works. The left figure represents the initial step of the algorithm, which calculates all the corner points. The right figure demonstrates the subsequent step, where points in-between the left and right are interpolated using the sine() function.}\label{fig:shade-algorithm}
\end{figure}

The algorithm behind the \texttt{sankey\_shade()} function for generating the proportional shaded data is illustrated in Figure \ref{fig:shade-algorithm}. It started by calculating the corner points for all of the shaded areas. Then, use the \texttt{sine()} function to interpolate the point between the left and right.

@@

\caption{\label{tab:unnamed-chunk-8}The arguments for the wallaby\_plot function.}
\centering
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{30em}}
\hline
Argument & Description\\
\hline
object & The animbook object returned from the prep function.\\
\hline
group\_palette & The vector of the palette used by the function to supply the color to each group.\\
\hline
shade\_palette & The vector of the palette used by the function to supply the color to the shaded area.\\
\hline
rendering & The choice of method used to create and display the plot, either gganimate or plotly.\\
\hline
subset & A character string specifying the variable used for subsetting the data. The "top" and "bottom" strings can also be used in this argument.\\
\hline
relation & The choice of relationship for the values to display on the plot, either "one\_many." or "many\_one."\\
\hline
total\_point & The number of points the users want for the wallaby plot. Default is NULL, which is the number of points equal to the original.\\
\hline
height & The proportion of the area occupied by the observations in the shaded areas.\\
\hline
width & The distance between the first and last observation in the animation.\\
\hline
size & The point size.\\
\hline
alpha & The opacity of the proportional shaded areas.\\
\hline
\end{tabular}
\end{table}

@@

In this section, the package usage will be demonstrated using the \texttt{cat\_change} dataset included in this package. This dataset is simulated data that has changed from category A to E between two time points. Since this data is already in the categorized data structure, the \texttt{anim\_prep\_cat()} function is utilized.

The output produced from this package is referred to as the \texttt{animbook} object. This contained a list of data and settings that hold prepared data for visualization, along with a list of settings used in the plot function.

\begin{verbatim}
animbook <- anim_prep_cat(cat_change, 
                          id = id, 
                          values = qnt, 
                          time = time, 
                          color = gp, 
                          time_dependent = FALSE)

head(animbook$data, 10)
\end{verbatim}

\begin{verbatim}
#> # A tibble: 10 x 5
#>    id     time qtile frame color
#>    <fct> <int> <dbl> <dbl> <fct>
#>  1 1      2020     5    36 X    
#>  2 1      2023     5    37 X    
#>  3 2      2020     5    49 X    
#>  4 2      2023     5    50 X    
#>  5 3      2020     5    48 X    
#>  6 3      2023     5    49 X    
#>  7 4      2020     5    50 X    
#>  8 4      2023     5    51 X    
#>  9 5      2020     5    43 X    
#> 10 5      2023     5    44 X
\end{verbatim}

@@

\begin{verbatim}
#> List of 2
#>  $ data    : tibble [400 x 5] (S3: tbl_df/tbl/data.frame)
#>   ..$ id   : Factor w/ 200 levels "1","2","3","4",..: 1 1 2 2 3 3 4 4 5 5 ...
#>   ..$ time : int [1:400] 2020 2023 2020 2023 2020 2023 2020 2023 2020 2023 ...
#>   ..$ qtile: num [1:400] 5 5 5 5 5 5 5 5 5 5 ...
#>   ..$ frame: num [1:400] 36 37 49 50 48 49 50 51 43 44 ...
#>   ..$ color: Factor w/ 2 levels "X","Y": 1 1 1 1 1 1 1 1 1 1 ...
#>  $ settings:List of 7
#>   ..$ gap           : num 0.1
#>   ..$ xbreaks       : int [1:2] 2020 2023
#>   ..$ label         : chr [1:5] "A" "B" "C" "D" ...
#>   ..$ order         : chr [1:5] "A" "B" "C" "D" ...
#>   ..$ time_dependent: logi FALSE
#>   ..$ runif_min     : num 1
#>   ..$ runif_max     : num 50
#>  - attr(*, "class")= chr "animbook"
\end{verbatim}

@@

\begin{verbatim}
p <- wallaby_plot(object = animbook,
             group_palette = RColorBrewer::brewer.pal(9, "Set1"),
             shade_palette = c("#737373", "#969696", "#BDBDBD",
                               "#D9D9D9","#D9D9D9","#D9D9D9"),
             rendering = "ggplot",
             subset = "top",
             relation = "one_many",
             total_point = NULL)

@@

\begin{verbatim}
voter <- anim_prep_cat(data = aeles,
                       id = id,
                       values = party,
                       time = year,
                       color = gender,
                       order = NULL,
                       time_dependent = FALSE)

p_voter <- wallaby_plot(object = voter,
                  group_palette = c("pink", "blue", "red"),
                  shade_palette = c("#737373", "#969696", "#BDBDBD",
                                    "#D9D9D9","#D9D9D9","#D9D9D9"),
                  rendering = "ggplot",
                  subset = "top",
                  relation = "one_many",
                  height = 1,
                  size = 2.5,
                  width = 100,
                  total_point = 1000)

>>>>

In designing the \pkg{animbook} package, a three-step structured approach was developed to create an animation. The initial step is to reformat the data into a categorized data structure, as seen in Figure \ref{fig:data-diagram}. The second stage is creating a \texttt{ggplot} object, which can be subsequently passed into the animation function. During the second stage, an internal function will turn the categorized data into an animated data structure. The final step involves transforming the \texttt{ggplot} to a \texttt{gganimate} object by integrating the animation settings. This three-step structure was implemented to ensure that users, regardless of their level of experience, can produce the animations with simplicity while retaining customization for more experienced users.

@@

The frame variable is assigned by sorting the data based on the ID and time using the \texttt{arrange()} function, followed by applying the \texttt{group\_by()} function on the ID, allowing the \texttt{row\_number()} function to be performed within each group. The functions mentioned in this paragraph are from the \CRANpkg{dplyr} package (Wickham et al. (2023)). These are done by the internal function for each of the plots provided in the packages.

@@

\caption{\label{tab:tbl-animprep}The argument in the anim\_prep for customized the data scaling.}
\centering
\begin{tabular}[t]{l|l}
\hline
Argument & Description\\
\hline
ncat & The number of groups or categories to create for scaling values.\\
\hline
breaks & A vector of breaks for creating bins.\\
\hline
group\_scaling & The column name that represents the grouping variable.\\
\hline
scaling & The scaling method to be used; "rank" or "absolute."\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:tbl-visual}The additional argument for customizing the aesthetics of the visualization.}
\centering
\begin{tabular}[t]{l|l}
\hline
Argument & Description\\
\hline
group & The column name that distinguishes between the values group.\\
\hline
label & A vector of labels to be used for the y-axis in the visualization.\\
\hline
time\_dependent & Logical. Should the visualization be time-dependent? Default is TRUE.\\
\hline
\end{tabular}
\end{table}

@@

Once the data is prepared. The next step is to create the \texttt{ggplot} object as a basis for the animation. There are three plots available in this package. Two of the plots could be used for the animation, and another plot is used as a static visualization. All of the plots have an internal function that converts the categorized data format into the animated structure for each plotting function.

@@

The \texttt{wallaby\_data()} function is responsible for performing data manipulation and formatting tasks on the original object. It includes creating additional data components for labeling and shading. This function also responds by interpolating the non-linear path. It performs this task by mapping the \texttt{sine()} function provided in this package to each observation using the \texttt{map()} function from the \CRANpkg{purrr} package (Wickham and Henry (2023)). The frame is recalculated using the same method mentioned in the data preprocessing section. The \texttt{proportional\_shade()} function is called to generate the proportional shaded data.

\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/sankey-shade-1} \includegraphics[width=0.5\linewidth]{figures/sankey-shade-2} \caption{The plot shows how the algorithm for the proportional\_shade() function works. The left figure represents the initial step of the algorithm, which calculates all the corner points. The right figure demonstrates the subsequent step, where points in-between the left and right are interpolated using the sine() function.}\label{fig:shade-algorithm}
\end{figure}

The algorithm behind the \texttt{proportional\_shade()} function for generating the proportional shaded data is illustrated in Figure \ref{fig:shade-algorithm}. It started by calculating the corner points for all of the shaded areas. Then, use the \texttt{sine()} function to interpolate the point between the left and right.

@@

\caption{\label{tab:unnamed-chunk-8}The arguments for the wallaby\_plot function.}
\centering
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{30em}}
\hline
Argument & Description\\
\hline
data & The categorized data returned from the prep function.\\
\hline
group\_palette & The vector of the palette used by the function to supply the color to each group.\\
\hline
shade\_palette & The vector of the palette used by the function to supply the color to the shaded area.\\
\hline
rendering & The choice of method used to create and display the plot, either gganimate or plotly.\\
\hline
subset & A character string specifying the variable used for subsetting the data. The "top" and "bottom" strings can also be used in this argument.\\
\hline
relation & The choice of relationship for the values to display on the plot, either "one\_many." or "many\_one."\\
\hline
total\_point & The number of points the users want for the wallaby plot. Default is NULL, which is the number of points equal to the original.\\
\hline
height & The proportion of the area occupied by the observations in the shaded areas.\\
\hline
width & The distance between the first and last observation in the animation.\\
\hline
size & The point size.\\
\hline
alpha & The opacity of the proportional shaded areas.\\
\hline
\end{tabular}
\end{table}

@@

In this section, the package usage will be demonstrated using the \texttt{cat\_change} dataset included in this package. This dataset is simulated data that has changed from category A to E between two time points. Since this data is already in the categorized data structure, the \texttt{anim\_prep\_cat()} function is utilized. The output produced from this package is referred to as the \texttt{categorized} data.

\begin{verbatim}
animbook <- anim_prep_cat(cat_change, 
                          id = id, 
                          values = qnt, 
                          time = time, 
                          group = gp)

head(animbook, 10)
\end{verbatim}

\begin{verbatim}
#> # A tibble: 10 x 5
#>    id     time qtile label group
#>    <fct> <int> <dbl> <chr> <fct>
#>  1 1      2020     5 A     X    
#>  2 2      2020     5 A     X    
#>  3 3      2020     5 A     X    
#>  4 4      2020     5 A     X    
#>  5 5      2020     5 A     X    
#>  6 6      2020     5 A     X    
#>  7 7      2020     5 A     X    
#>  8 8      2020     5 A     X    
#>  9 9      2020     5 A     X    
#> 10 10     2020     5 A     X
\end{verbatim}

@@

\begin{verbatim}
#> tibble[,5] (S3: tbl_df/tbl/data.frame/categorized)
#>  $ id   : Factor w/ 200 levels "1","2","3","4",..: 1 2 3 4 5 6 7 8 9 10 ...
#>  $ time : int [1:400] 2020 2020 2020 2020 2020 2020 2020 2020 2020 2020 ...
#>  $ qtile: num [1:400] 5 5 5 5 5 5 5 5 5 5 ...
#>  $ label: chr [1:400] "A" "A" "A" "A" ...
#>  $ group: Factor w/ 2 levels "X","Y": 1 1 1 1 1 1 1 1 1 1 ...
\end{verbatim}

@@

\begin{verbatim}
p <- wallaby_plot(data = animbook,
             group_palette = RColorBrewer::brewer.pal(9, "Set1"),
             shade_palette = c("#737373", "#969696", "#BDBDBD",
                               "#D9D9D9","#D9D9D9","#D9D9D9"),
             rendering = "gganimate",
             subset = "top",
             relation = "one_many",
             total_point = NULL)

@@

\begin{verbatim}
voter <- anim_prep_cat(data = aeles,
                       id = id,
                       values = party,
                       time = year,
                       group = gender,
                       order = NULL)

p_voter <- wallaby_plot(data = voter,
                  group_palette = c("pink", "blue", "red"),
                  shade_palette = c("#737373", "#969696", "#BDBDBD",
                                    "#D9D9D9","#D9D9D9","#D9D9D9"),
                  rendering = "gganimate",
                  subset = "top",
                  relation = "one_many",
                  height = 1,
                  size = 2.5,
                  width = 100,
                  total_point = 1000)
